Group Theory in Physics (Course Note)

### 2016-11-15 Lecture note - Group Theory in Physics/group theory in physics.tex
Conjugacy classes and regular representation in Group theory.
Finished another proof for "conjugacy class"="irrep" in group representation theory.
Adding point group discussion. Introducing the dihedral and rotation group.
New notes for algebraic theory of representation.

\maketitle
\abstract{
This is our course note for the course about group theoy, with its
application in physics.
}
\tableofcontents
\begin{enumerate}
    \item Combine group theory with you research.
    \item Mid-term and final.
    \item No homework, cause it is already graduate level.
    \item Professor Fei Ye. (Phone: 88018229, 228 in Research building 2), 
        T.A. Zhe Zhang. (110 Research building 2).
\end{enumerate}

@@

\section{20161031}
\label{sec:20161031}
\input{sections/20161031}
\section{Anchor}
\label{sec:Anchor}
\begin{thebibliography}{1}
    \bibitem{book} Zhongqi Ma, Group Theory in Physics
    \bibitem{math.se_1_lenz_vector} \href{physics.stackexchange.com/questions/18088/what-symmetry-causes-the-runge-lenz-vector-to-be-conserved}{What symmetry causes the Runge-Lenz vector to be conserved?}
    \bibitem{Ludeling} Lecture Notes for physics751: Group Theory (for
    Physicists), by C Ludeling.
    \href{http://www.th.physik.uni-bonn.de/nilles/people/luedeling/grouptheory/data/grouptheorynotes.pdf}{Link}
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}

>>>>

\maketitle
\abstract{
This is our course note for the course about group theoy, with its
application in physics.
}
\tableofcontents
\paragraph{Information about the class}
\begin{enumerate}
    \item Final project: Combine group theory with you research.
    \item Mid-term and final exam.
    \item No homework, cause it is already graduate level.
    \item Professor Fei Ye. (Phone: 88018229, 228 in Research building 2), 
        T.A. Zhe Zhang. (110 Research building 2).
\end{enumerate}

@@

\section{20161031}
\label{sec:20161031}
\input{sections/20161031}

\section{20161107}
\label{sec:20161107}
\input{sections/20161107}

\section{Point Group}
\label{sec:Point-Group}
\input{sections/point-group}

\section{Representationas and Semisimplicity}
\label{sec:Representationas-and-Semisimplicity}

Now I digress to read chapter 18 of the book \cite{lang-algebra}.

Let $R$ be a commutative ring and $G$ a group. Let $E$ be a
$R$-module.

For convenience, given a representation $\rho:G\to \Aut_R(E)$, let
$\sigma\ni G$, $x\in E$, we write $\sigma x$ instead of
$\rho(\sigma)x$.

An $R$-module $E$, together with a representation $\rho$, will be
called a \textbf{$G$-module}\nomenclature{$G$-module}{\nomrefpage} or
\textbf{$G$-space}\nomenclature{$G$-space}{\nomrefpage}, or also a
\textbf{$(G,R)$-module}\nomenclature{$(G,R)$-module}{\nomrefpage}.

If $E$, $F$ are $G$-modules, we define a
\textbf{$G$-homomorphism}\nomenclature{$G$-homomorphism}{\nomrefpage}
$f:E\to F$ as a $R$-linear map such that $f(\sigma x) = \sigma f(x)$
for all $x\in G$ and $\sigma\in G$.

We make $R$ into a $G$-module by making $G$ act trivially on $R$, i.e.
we have a trivial representation on $R$, itself regarded as a
$R$-module.

\section{Anchor}
\label{sec:Anchor}
\begin{thebibliography}{1}
    \bibitem{book} Zhongqi Ma, Group Theory in Physics
    \bibitem{math.se_1_lenz_vector} \href{physics.stackexchange.com/questions/18088/what-symmetry-causes-the-runge-lenz-vector-to-be-conserved}{What symmetry causes the Runge-Lenz vector to be conserved?}
    \bibitem{Ludeling} Lecture Notes for physics751: Group Theory (for
    Physicists), by C Ludeling.
    \href{http://www.th.physik.uni-bonn.de/nilles/people/luedeling/grouptheory/data/grouptheorynotes.pdf}{Link}
    \bibitem{lang-algebra} Serge Lang. Algebra. Revised 3rd. Springer.
\end{thebibliography}
\printnomenclature
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}

### 2016-11-15 Lecture note - Group Theory in Physics/sections/point-group.tex
New notes for algebraic theory of representation.

The general procedure is to use the invariant subgroup of index $2$
mentioned above to get the $1$-dimensional representations. And use
the relation $\sum_j m_j^2 = |D_n|$ and $\sum_j 1 = $ number of
conjugacy class to "guess" the dimensional of remaining irreducible
representations. Finally use subgroup $C_n$ to construct those
representations (induced representation) and reduce these into
irreducible ones. But how to reduce them into irreducible one is
actually the most difficult part, in my opinion.

I present only vague discussion below. For details, please to ...
actually I don't know which book. The textbook \cite{book} is too
vague too. Perhaps I should find some more detailed books.
% TODO Add better books.

@@

There is two $1$-dimensional irreducible representations, formed using
the invariant subgroup $C_{2n'+1}$. Since $4n'+2 = 2^2 n'+2$, one
guesses that the remaining ones are $n$ $2$-dimensional irreducible
representation. They are constructed (construction process could be
found in \cite{book}) to be (notice $D_{2n'+1}$ is generated by
$T_{2n'+1}$, the rotation $2\pi/(2n'+1)$, and $C'_2$, any one of the
$2$-fold reflection):
\begin{align}
    \bar{D}^{E_j}(C_{n}) &=\left( \begin{array}{cc}
    \cos(\frac{2j\pi}{n}) & -\sin(\frac{2j\pi}{n}) \\
    \sin(\frac{2j\pi}{n}) & \cos(\frac{2j\pi}{n}) \\
        \end{array} \right) \\
    \bar{D}^{E_j}(C'_{2}) &= \left( \begin{array}{cc}
                 1 & 0 \\
                 0 & -1 \\ 
                 \end{array} \right)
\end{align}
where $j=1,2,\cdots,n$.

>>>>

The general procedure is to use the invariant subgroup of index $2$
mentioned above to get the $1$-dimensional representations. And use
the relation $\sum_j m_j^2 = |D_n|$ and $\sum_j 1 = $ number of
conjugacy class to get the dimensional of remaining irreducible
representations. Finally use subgroup $C_n$ to construct those
representations (induced representation) and reduce these into
irreducible ones. But how to reduce them into irreducible one is
actually the most difficult part, in my opinion.

I present only vague discussion below. For details, please to ...
actually I don't know which book. The textbook \cite{book} is too
concise. Perhaps I should find some more detailed books.
% TODO Add better books.

@@

There is two $1$-dimensional irreducible representations, formed using
the invariant subgroup $C_{2n'+1}$. Since $4n'+2 = 2^2 n'+2$, one
guesses\footnote{Actually, the remaing ones can satisfy our condition
    only if they are $2$-dimensional. A $1$-dimensional would means an
    additional invarian subgroup of index $2$, which is not true. A
    $3$-dimensional would demand an additional $1$-dimensional
    representation to balance and get a even number. So
$2$-dimensional is the only possibility.}, that the remaining ones are
$n$ $2$-dimensional irreducible representations. They are constructed
(construction process could be found in \cite{book}) to be (notice
$D_{2n'+1}$ is generated by $T_{2n'+1}$, the rotation
$2\pi/(2n'+1)$, and $C'_2$, any one of the $2$-fold reflection):
\begin{align}
    \bar{D}^{E_j}(C_{n}) &=\left( \begin{array}{cc}
    \cos(\frac{2j\pi}{n}) & -\sin(\frac{2j\pi}{n}) \\
    \sin(\frac{2j\pi}{n}) & \cos(\frac{2j\pi}{n}) \\
        \end{array} \right) \\
    \bar{D}^{E_j}(C'_{2}) &= \left( \begin{array}{cc}
                 1 & 0 \\
                 0 & -1 \\ 
                 \end{array} \right)
\end{align}
where $j=1,2,\cdots,n$.
